---
title: "Programa de Prevención y Control de Arbovirosis | 2026 | Sonora"
format: 
  dashboard:
    nav-buttons: [twitter, github, facebook]
    logo: https://framework-gb.cdn.gob.mx/gobmx/img/logo_blanco.svg
theme: custom.scss
---

```{r setup}
library(sf)
library(magrittr)
library(mapgl)
library(fishualize)
load("/Users/fdzul/Documents/geocoding_mx_2026_dengue/4.RData/dengue_mx_2026.RData")
source("~/Library/CloudStorage/Dropbox/r_developments/r_dashboards/netlify/1.2.persisten_dengue_transmission_2026/persistent_dengue_transmission_10_durango/mp_heatmap.R")
source("~/Library/CloudStorage/Dropbox/r_developments/r_dashboards/netlify/1.2.persisten_dengue_transmission_2026/persistent_dengue_transmission_10_durango/mp_transmission_chains.R")
source("~/Library/CloudStorage/Dropbox/r_developments/r_dashboards/netlify/1.2.persisten_dengue_transmission_2026/persistent_dengue_transmission_10_durango/mp_dengue_cases.R")
```

```{=html}
<style>
* { box-sizing: border-box; }

.responsive-grid {
  display: grid;
  grid-template-columns: 1fr 1fr 1fr;
  gap: 25px;
  margin: 20px 0;
  max-height: 85vh;
  overflow-y: auto;
  padding: 15px;
  border: 1px solid #e0e0e0;
  border-radius: 12px;
  background-color: #fafafa;
  align-items: start;
}

.columna {
  display: flex;
  flex-direction: column;
  gap: 20px;
}

.column-header {
  font-weight: bold;
  font-size: 1rem;
  text-align: center;
  padding: 8px;
  background-color: #772953;
  color: white;
  border-radius: 6px;
  margin: 0;
}

.grid-card {
  border: 1px solid #e0e0e0;
  padding: 15px;
  border-radius: 12px;
  background: white;
  transition: transform 0.2s, box-shadow 0.2s;
  box-shadow: 0 4px 6px rgba(0,0,0,0.05);
}

.grid-card:hover {
  transform: translateY(-3px);
  box-shadow: 0 8px 15px rgba(0,0,0,0.1);
}

.grid-card .html-widget,
.grid-card iframe {
  width: 100% !important;
  height: 400px !important;
  border-radius: 8px;
}

@media (max-width: 768px) {
  .responsive-grid {
    grid-template-columns: 1fr;
    max-height: 70vh;
    gap: 15px;
    padding: 10px;
  }
  .grid-card .html-widget,
  .grid-card iframe {
    height: 300px !important;
  }
}
</style>
```


# [**Mapas de Calor**]{style="font-size:8pt"}

::: panel-tabset

### [**Cajeme**]{style="color:#990000"}

::::::::::::::::::::: responsive-grid

<div class="columna">
<div class="column-header">Enero-Marzo</div>

::: grid-card
```{r heatmap_enero_marzo_cajeme, echo=FALSE}
z |>
    dplyr::filter(IDE_EDA_ANO <= 12 | IDE_EDA_ANO >= 65) |>
    dplyr::mutate(onset = FEC_INI_SIGNOS_SINT) |>
    dplyr::select(ANO, long, lat, onset) |>
    dplyr::mutate(x = long, y = lat) |>
    sf::st_as_sf(coords = c("x", "y"), crs = 4326) |>
    dplyr::mutate(month = lubridate::month(onset)) |>
    dplyr::filter(month %in% c(1:3)) |>
    mp_heatmap(yrs = 2026, cve_edo = "26", locality = "Ciudad Obregon") |>
    htmlwidgets::onRender("
        function(el, x) {
            setTimeout(function() {
                if (el._map) el._map.resize();
                window.dispatchEvent(new Event('resize'));
            }, 200);
        }
    ")
```
:::

</div>

<div class="columna">
<div class="column-header">Abril</div>

::: grid-card
```{r heatmap_abril_cajeme, echo=FALSE}
z |>
    dplyr::filter(IDE_EDA_ANO <= 12 | IDE_EDA_ANO >= 65) |>
    dplyr::mutate(onset = FEC_INI_SIGNOS_SINT) |>
    dplyr::select(ANO, long, lat, onset) |>
    dplyr::mutate(x = long, y = lat) |>
    sf::st_as_sf(coords = c("x", "y"), crs = 4326) |>
    dplyr::mutate(month = lubridate::month(onset)) |>
    dplyr::filter(month %in% c(4)) |>
    mp_heatmap(yrs = 2026, cve_edo = "26", locality = "Ciudad Obregon") |>
    htmlwidgets::onRender("
        function(el, x) {
            setTimeout(function() {
                if (el._map) el._map.resize();
                window.dispatchEvent(new Event('resize'));
            }, 200);
        }
    ")
```
:::

</div>

<div class="columna">
<div class="column-header">Enero-Abril</div>

::: grid-card
```{r heatmap_enero_abril_cajeme, echo=FALSE}
z |>
    dplyr::filter(IDE_EDA_ANO <= 12 | IDE_EDA_ANO >= 65) |>
    dplyr::mutate(onset = FEC_INI_SIGNOS_SINT) |>
    dplyr::select(ANO, long, lat, onset) |>
    dplyr::mutate(x = long, y = lat) |>
    sf::st_as_sf(coords = c("x", "y"), crs = 4326) |>
    mp_heatmap(yrs = 2026, cve_edo = "26", locality = "Ciudad Obregon")
```
:::

</div>

:::::::::::::::::::::


### [**Hermosillo**]{style="color:#990000"}

::::::::::::::::::::: responsive-grid

<div class="columna">
<div class="column-header">Enero-Marzo</div>

::: grid-card
```{r heatmap_enero_marzo_hermosillo, echo=FALSE}
z |>
    dplyr::filter(IDE_EDA_ANO <= 12 | IDE_EDA_ANO >= 65) |>
    dplyr::mutate(onset = FEC_INI_SIGNOS_SINT) |>
    dplyr::select(ANO, long, lat, onset) |>
    dplyr::mutate(x = long, y = lat) |>
    sf::st_as_sf(coords = c("x", "y"), crs = 4326) |>
    dplyr::mutate(month = lubridate::month(onset)) |>
    dplyr::filter(month %in% c(1:3)) |>
    mp_heatmap(yrs = 2026, cve_edo = "26", locality = "Hermosillo")
```
:::

</div>

<div class="columna">
<div class="column-header">Abril</div>

::: grid-card
```{r heatmap_abril_hermosillo, echo=FALSE}
z |>
    dplyr::filter(IDE_EDA_ANO <= 12 | IDE_EDA_ANO >= 65) |>
    dplyr::mutate(onset = FEC_INI_SIGNOS_SINT) |>
    dplyr::select(ANO, long, lat, onset) |>
    dplyr::mutate(x = long, y = lat) |>
    sf::st_as_sf(coords = c("x", "y"), crs = 4326) |>
    dplyr::mutate(month = lubridate::month(onset)) |>
    dplyr::filter(month %in% c(4)) |>
    mp_heatmap(yrs = 2026, cve_edo = "26", locality = "Hermosillo")
```
:::

</div>

<div class="columna">
<div class="column-header">Enero-Abril</div>

::: grid-card
```{r heatmap_enero_abril_hermosillo, echo=FALSE}
z |>
    dplyr::filter(IDE_EDA_ANO <= 12 | IDE_EDA_ANO >= 65) |>
    dplyr::mutate(onset = FEC_INI_SIGNOS_SINT) |>
    dplyr::select(ANO, long, lat, onset) |>
    dplyr::mutate(x = long, y = lat) |>
    sf::st_as_sf(coords = c("x", "y"), crs = 4326) |>
    mp_heatmap(yrs = 2026, cve_edo = "26", locality = "Hermosillo")
```
:::

</div>

:::::::::::::::::::::


### [**Guaymas**]{style="color:#990000"}

::::::::::::::::::::: responsive-grid

<div class="columna">
<div class="column-header">Enero-Marzo</div>

::: grid-card
```{r heatmap_enero_marzo_guaymas, echo=FALSE}
z |>
    dplyr::filter(IDE_EDA_ANO <= 12 | IDE_EDA_ANO >= 65) |>
    dplyr::mutate(onset = FEC_INI_SIGNOS_SINT) |>
    dplyr::select(ANO, long, lat, onset) |>
    dplyr::mutate(x = long, y = lat) |>
    sf::st_as_sf(coords = c("x", "y"), crs = 4326) |>
    dplyr::mutate(month = lubridate::month(onset)) |>
    dplyr::filter(month %in% c(1:3)) |>
    mp_heatmap(yrs = 2026, cve_edo = "26", locality = "Heroica Guaymas")
```
:::

</div>

<div class="columna">
<div class="column-header">Abril</div>

::: grid-card
```{r heatmap_abril_guaymas, echo=FALSE}
z |>
    dplyr::filter(IDE_EDA_ANO <= 12 | IDE_EDA_ANO >= 65) |>
    dplyr::mutate(onset = FEC_INI_SIGNOS_SINT) |>
    dplyr::select(ANO, long, lat, onset) |>
    dplyr::mutate(x = long, y = lat) |>
    sf::st_as_sf(coords = c("x", "y"), crs = 4326) |>
    dplyr::mutate(month = lubridate::month(onset)) |>
    dplyr::filter(month %in% c(4)) |>
    mp_heatmap(yrs = 2026, cve_edo = "26", locality = "Heroica Guaymas")
```
:::

</div>

<div class="columna">
<div class="column-header">Enero-Abril</div>

::: grid-card
```{r heatmap_enero_abril_guaymas, echo=FALSE}
z |>
    dplyr::filter(IDE_EDA_ANO <= 12 | IDE_EDA_ANO >= 65) |>
    dplyr::mutate(onset = FEC_INI_SIGNOS_SINT) |>
    dplyr::select(ANO, long, lat, onset) |>
    dplyr::mutate(x = long, y = lat) |>
    sf::st_as_sf(coords = c("x", "y"), crs = 4326) |>
    mp_heatmap(yrs = 2026, cve_edo = "26", locality = "Heroica Guaymas")
```
:::

</div>

:::::::::::::::::::::


### [**Empalme**]{style="color:#990000"}

::::::::::::::::::::: responsive-grid

<div class="columna">
<div class="column-header">Enero-Marzo</div>

::: grid-card
```{r heatmap_enero_marzo_empalme, echo=FALSE}
z |>
    dplyr::filter(IDE_EDA_ANO <= 12 | IDE_EDA_ANO >= 65) |>
    dplyr::mutate(onset = FEC_INI_SIGNOS_SINT) |>
    dplyr::select(ANO, long, lat, onset) |>
    dplyr::mutate(x = long, y = lat) |>
    sf::st_as_sf(coords = c("x", "y"), crs = 4326) |>
    dplyr::mutate(month = lubridate::month(onset)) |>
    dplyr::filter(month %in% c(1:3)) |>
    mp_heatmap(yrs = 2026, cve_edo = "26", locality = "Empalme")
```
:::

</div>

<div class="columna">
<div class="column-header">Abril</div>

::: grid-card
```{r heatmap_abril_empalme, echo=FALSE}
z |>
    dplyr::filter(IDE_EDA_ANO <= 12 | IDE_EDA_ANO >= 65) |>
    dplyr::mutate(onset = FEC_INI_SIGNOS_SINT) |>
    dplyr::select(ANO, long, lat, onset) |>
    dplyr::mutate(x = long, y = lat) |>
    sf::st_as_sf(coords = c("x", "y"), crs = 4326) |>
    dplyr::mutate(month = lubridate::month(onset)) |>
    dplyr::filter(month %in% c(4)) |>
    mp_heatmap(yrs = 2026, cve_edo = "26", locality = "Empalme")
```
:::

</div>

<div class="columna">
<div class="column-header">Enero-Abril</div>

::: grid-card
```{r heatmap_enero_abril_empalme, echo=FALSE}
z |>
    dplyr::filter(IDE_EDA_ANO <= 12 | IDE_EDA_ANO >= 65) |>
    dplyr::mutate(onset = FEC_INI_SIGNOS_SINT) |>
    dplyr::select(ANO, long, lat, onset) |>
    dplyr::mutate(x = long, y = lat) |>
    sf::st_as_sf(coords = c("x", "y"), crs = 4326) |>
    mp_heatmap(yrs = 2026, cve_edo = "26", locality = "Empalme")
```
:::

</div>

:::::::::::::::::::::


### [**Navojoa**]{style="color:#990000"}

::::::::::::::::::::: responsive-grid

<div class="columna">
<div class="column-header">Enero-Marzo</div>

::: grid-card
```{r heatmap_enero_marzo_navojoa, echo=FALSE}
z |>
    dplyr::filter(IDE_EDA_ANO <= 12 | IDE_EDA_ANO >= 65) |>
    dplyr::mutate(onset = FEC_INI_SIGNOS_SINT) |>
    dplyr::select(ANO, long, lat, onset) |>
    dplyr::mutate(x = long, y = lat) |>
    sf::st_as_sf(coords = c("x", "y"), crs = 4326) |>
    dplyr::mutate(month = lubridate::month(onset)) |>
    dplyr::filter(month %in% c(1:3)) |>
    mp_heatmap(yrs = 2026, cve_edo = "26", locality = "Navojoa")
```
:::

</div>

<div class="columna">
<div class="column-header">Abril</div>

::: grid-card
```{r heatmap_abril_navojoa, echo=FALSE}
z |>
    dplyr::filter(IDE_EDA_ANO <= 12 | IDE_EDA_ANO >= 65) |>
    dplyr::mutate(onset = FEC_INI_SIGNOS_SINT) |>
    dplyr::select(ANO, long, lat, onset) |>
    dplyr::mutate(x = long, y = lat) |>
    sf::st_as_sf(coords = c("x", "y"), crs = 4326) |>
    dplyr::mutate(month = lubridate::month(onset)) |>
    dplyr::filter(month %in% c(4)) |>
    mp_heatmap(yrs = 2026, cve_edo = "26", locality = "Navojoa")
```
:::

</div>

<div class="columna">
<div class="column-header">Enero-Abril</div>

::: grid-card
```{r heatmap_enero_abril_navojoa, echo=FALSE}
z |>
    dplyr::filter(IDE_EDA_ANO <= 12 | IDE_EDA_ANO >= 65) |>
    dplyr::mutate(onset = FEC_INI_SIGNOS_SINT) |>
    dplyr::select(ANO, long, lat, onset) |>
    dplyr::mutate(x = long, y = lat) |>
    sf::st_as_sf(coords = c("x", "y"), crs = 4326) |>
    mp_heatmap(yrs = 2026, cve_edo = "26", locality = "Navojoa")
```
:::

</div>

:::::::::::::::::::::

:::


# [**Cadenas de Transmisión**]{style="font-size:8pt"}

::: panel-tabset

### [**Hermosillo**]{style="color:#990000"}

::::::::::::::::::::: responsive-grid

<div class="columna">
<div class="column-header">Distribución de Casos</div>

::: grid-card
```{r distribucion_hermosillo, echo=FALSE}
z |>
    dplyr::filter(IDE_EDA_ANO <= 12 | IDE_EDA_ANO >= 65) |>
    dplyr::mutate(onset = FEC_INI_SIGNOS_SINT) |>
    dplyr::select(ANO, long, lat, onset) |>
    dplyr::mutate(x = long, y = lat) |>
    sf::st_as_sf(coords = c("x", "y"), crs = 4326) |>
    mp_dengue_cases(yrs = 2026, locality = "Hermosillo", cve_edo = "26")
```
:::

</div>

<div class="columna">
<div class="column-header">Cadenas de Transmisión</div>

::: grid-card
```{r chains_hermosillo, echo=FALSE}
z |>
    dplyr::filter(IDE_EDA_ANO <= 12 | IDE_EDA_ANO >= 65) |>
    dplyr::mutate(onset = FEC_INI_SIGNOS_SINT) |>
    dplyr::select(ANO, long, lat, onset) |>
    dplyr::mutate(x = long, y = lat) |>
    sf::st_as_sf(coords = c("x", "y"), crs = 4326) |>
    mp_transmission_chains(yrs = 2026, locality = "Hermosillo", cve_edo = "26")
```
:::

</div>

<div class="columna">
<div class="column-header">Mapa de Calor</div>

::: grid-card
```{r heatmap_cadenas_hermosillo, echo=FALSE}
z |>
    dplyr::filter(IDE_EDA_ANO <= 12 | IDE_EDA_ANO >= 65) |>
    dplyr::mutate(onset = FEC_INI_SIGNOS_SINT) |>
    dplyr::select(ANO, long, lat, onset) |>
    dplyr::mutate(x = long, y = lat) |>
    sf::st_as_sf(coords = c("x", "y"), crs = 4326) |>
    mp_heatmap(yrs = 2026, cve_edo = "26", locality = "Hermosillo")
```
:::

</div>

:::::::::::::::::::::

:::
